\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing economic inequality with tax data - Switzerland from 1945 to 2010}

### 2014-05-27 paper/subdocuments/data_and_methods.Rnw
minor changes data and method section

\section{ Data, Measurement concepts,and Methods}

Studies on inequality have to adress several thorny challenges. It starts with answering three crucial questions: First of all one has to define, which concepts should to be looked at. This refers to answering the question about inequality of what. Secondly one has to be clear about the unit of analysis. This refers to answering the question about inequality among whom. Thirdly, one has to choose an appropriate measure of inequality. All this questions are ideally answered considering theory and a given research question. Often it has to be answered in context of a given dataset. Based on a review on the literature about the measurement concepts in an ideal world, we discuss  the advantages and shortcomings of tax data compared to other data sources - namely survey data. Afterwards we describe the tax data published by the Swiss Federal Tax Administration highlighting the important aspects, which have to be considered working with FTA-Data. In the last section, we describe the methods and corrections we used to construct the time-series of inequality-measure for income for Switzerland.

% Allenfalls mit dem Beschrieb der Daten beginnen und bereits stärker auf die Möglichkeiten und Schwierigkeiten von Steuerdaten eingehen

@@

A lot of studies on inequality focus on income inequality solely. However, recent activities emphasize the need of a broader conceptualization. A recent publication from the oecd (2013) condense these activities into the ICW framework (income, consumption and wealth), which is ment to be an internationally agreed framework of micro-level statistics.\footnote{Harmonisation with other international standards was an important objective that guided the work of the Expert Group in developing the ICW Framewrok presented in this publication. Considered main stadards were the System of National Accounts (SNA 2008), the canberra Group Handbook on Household Income Statistics (United Nations 2011), the final report of the 17th International Conference of Labour Statisticians (ILO 2004) and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing (UNECE & Eurostat 2006)} According to the framework it is best to look at income, consumption and wealth as three separate but interrelated dimensions of people's economic well-being. It is recomended to look at the distribution of all three distributions together to gain policy relevant insight. Some households with low income, for example, may report adequate levels of consumption expenditure or welth holdings, or vice-versa. But it is also stated (OED 2013:18):"[...] integrated analysis at the household level has significant data requirements that go beyond the measurement efforts currently undertaken in most countries."\footnote{The Luxembourg Wealth Study Database is currently facing this shortcomings by colecting and provding a database following ths broader concept of economic well-being. http://www.lisdatacenter.org/our-data/lws-database/} \\

@@

% Eine Grafik/Tabelle machen?
% Mit den Brülhart-Daten hätte man noch mehr...

The assessment of income inequality is influenced by the conceptual definition of the income itself. Marketincome or  disposable income for example differ by substantial meaning and be the expected degree of inequality. Therefore the awarness of the analyzed concept is crucial. Terminology can sligthly differ, while common concepts can be identified (for detailed discussion see: OECD (2013:44), Canberra group (2011:24)). Figure xy shows a stylized framework, which includes a distinction of common income sources \footnote{Income from production of household services for own consumption is excluded because this income is hard to measure and not covered in the FTA tax data} and shows the central steps of redistribution. Within this framework common income definitions are situated. These definitions are contrasted with income concepts, which can be derived out of the FTA Tax data.

@@

The central income according to tax data is the taxable income. It includes all reported incomes (income from employment, income from property and received transfers\footnote{Mean-tested benefits are not taxed and therefore not included in Tax data. Income for low income groups are therefore underestimated.}) minus several deductions. It is therefore neither a pre-transfer income  nor a post-transfer income measure. It's rather something inbetween. As the FTA tax statistitcs include some but not all deductions it is possible to calculate a sort of "total income". As some deducations can be interpreted as compulsary expenses similar to taxes to step towards the Total income is a step away from the income, which can be used for consumption. Similiar when calculating the disposable income out of the taxable income through accouting the reported federal taxes, this is a step towards the income, which is left in the basket for consumption (disposable income). Again it is not a "pure" disposable income, because cantonal, municipal taxes and taxes from churches are missing.   \\

@@

%

@@

>>>>

\section{Data, Measurement concepts,and Methods}

Studies on inequality have to adress several thorny challenges. It starts with answering three crucial questions: First of all one has to define, which concepts should to be looked at. This refers to answering the question about inequality of what. Secondly one has to be clear about the unit of analysis. This refers to answering the question about inequality among whom. Thirdly, one has to choose an appropriate measure of inequality. All this questions are ideally answered considering theory and a given research question. Often it has to be answered in context of a given dataset. Therefore we start this section with a description of the FTA-Tax Data. Based on a review on the literature about the measurement concepts in an ideal world, we discuss  the advantages and shortcomings of tax data compared to other data sources - namely survey data. Along with the despriction we explain the methods and thechnics we use to construct time-series of inequality-measure for income for Switzerland, which we will present in the result sections. By doing this, we demonstrate the suitability of the FTA Tax Data to assess the development of inequality in Switzerland

\subsection{Tax statistics in Swizterland}

@@

Most studies on inequality focus on income inequality solely. However, recent activities emphasize the need of a broader conceptualization. A recent publication from the oecd (2013) condense these ideas into the ICW framework (income, consumption and wealth), which is ment to be an internationally agreed framework on micro-level statistics.\footnote{Harmonisation with other international standards was an important objective that guided the work of the Expert Group in developing the ICW Framewrok presented in this publication. Considered main stadards were the System of National Accounts (SNA 2008), the canberra Group Handbook on Household Income Statistics (United Nations 2011), the final report of the 17th International Conference of Labour Statisticians (ILO 2004) and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing (UNECE & Eurostat 2006)} According to the framework it is best to look at income, consumption and wealth as three separate but interrelated dimensions of people's economic well-being. To gain policy relevant insight, it is recomended to look at the distribution of all three distributions simultaneously. Some households with low income, for example, may report adequate levels of consumption expenditure or welth holdings, or vice-versa. But it is also stated (OED 2013:18):"[...] integrated analysis at the household level has significant data requirements that go beyond the measurement efforts currently undertaken in most countries."\footnote{The Luxembourg Wealth Study Database is currently facing this shortcomings by colecting and provding a database following ths broader concept of economic well-being. http://www.lisdatacenter.org/our-data/lws-database/} \\

@@

The assessment of income inequality is influenced by the definition of the income itself. Marketincome or disposable income for example differ by substantial meaning and by the expected degree of inequality. Therefore the awarness of the analyzed concept is crucial. Terminology can sligthly differ, while common concepts can be identified (for detailed discussion see: OECD (2013:44), Canberra group (2011:24)). Figure xy shows a stylized framework, which includes a distinction of common income sources \footnote{Income from production of household services for own consumption is excluded because this income is hard to measure and not covered in the FTA tax data} and shows the central steps of redistribution. Within this framework common income definitions are situated. These definitions are contrasted with income concepts, which can be derived out of the FTA Tax data.

@@

% Quelle zu Abzügen:
% http://www.bilanz.ch/invest/steuern/netto-oder-reineinkommen
% Die rechte Seite mit den Steuergrössen könnte ausführlicher beschrieben werden
% Beispielsweise könnte man zeigen was Nettoeinkommen, Reineinkommen und steuerbares Einkommen sind
% Allerdings steht uns nur Reineinkommen und steuerbares Einkommen zur Verfügung

The central income according to tax data is the taxable income. It includes all reported incomes (income from employment, income from property and received transfers\footnote{Mean-tested benefits are not taxed and therefore not included in Tax data. Income for low income groups are therefore underestimated.}) minus several deductions. It is therefore neither a pre-transfer income  nor a post-transfer income measure. It's rather something inbetween. As the FTA tax statistitcs include some but not all deductions\footnote{The difference between the real Total income and the taxable income are deductions. This includes: professional expenses, travel expenses, interest on debt, Alimonies, Training costs, two-earner deduction,Party Contributions, Payment into pillar 3a, purchases in the pension fund and sideline deductions} it is possible to calculate a sort of "total income". As some deducations can be interpreted as compulsary expenses similar to taxes the step towards Total income is a step away from the income, which can be used for consumption. Similiar when calculating the disposable income out of the taxable income through accouting the reported federal taxes, this is a step towards the income, which is left in the basket for consumption (disposable income). Again it is not a "pure" disposable income, because cantonal, municipal taxes and taxes from churches are missing.   \\

% Die Deductions sind noch nicht gänzlich sauber abgebildet. Auf der Bilanz-Seite steht, vom steuerbaren Einkommen kommt man, indem vom Reineinkommen die Sozialabzüge(Kinder- und Fremdbetreung) sowie der Abzug für unterstützungsbedürftige Personen abzogen werden. 
% Auf den ESTV-Tabellen sind jedoch noch mehr Abzüge abgebildet. Was soll/muss/können wir nun abziehen? Bin verwirrt...

@@

% letzter Abschnitt muss noch auf die Brülhart-Daten hinweisen. Damit können wir eine äquivalenzeinkommen betrachten.
% Auch müsste in diesem Abschnitt auf die Probleme der Haushaltsunterscheidung eingegangen werden (oder ist im Abschnitt Population coverage? Dann müsste obiger Punkt auch da erwähnt werden.)

% Streichen
=======
% 
>>>>>>> d21c323... completed subsection "concepts on measuring economic resources" and "defining income"

@@

% Alter Textbaustein zu Vermögen
%\emph{Defining wealth}
%While the apropiateness of the conceptualization of income is widely discuded, there is no agreed definition of personal wealth yet and the appropriate methods of valuation are not always clear. Wealth can be defined as the marketable value of financial assets plus non-financial assets (housing and land) less debts (Credit Suisse 2011:5).
